---
title: "R for Data Science"
subtitle: "Chapter 11: Data Import"
author: "Luke Morris"
institute: "@LukeMorris"
date: " `r Sys.Date()` "
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css","custom.css"]
    includes:
      after_body: insert-logo.html
---


```{r xaringan-themer, include = FALSE}
library(tidyverse)
library(xaringanthemer)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

## Importing your own data

- The tidyverse comes with handy tools to get your own data into R.

- The most common function is `read_csv()`, the focus of this chapter.

- There's also... 
  - `read_csv2()` if you're ever dealing with numbers written in non-U.S. style
    (thousands and millions split up by periods, semicolon as delimiter)
  
  - `read_tsv()` if your columns are split up by tabs instead of columns
  
  - `read_delim()`, which lets you set your own delimiter to split columns
  
---

## Let's pull some data in!

```{r}
trucks <- read_csv("trucks.csv")

trucks
```

---

## A little extra help

If your file doesn't get straight to the point, there are ways to point R to the 
beginning of the relevant data: `skip =` and `comment =`

```{r}
read_csv("The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3", skip = 2)

read_csv("# A comment I want to skip
  x,y,z
  1,2,3", comment = "#")
```

---

## A little extra help

No column names? No problem! You can tell R they don't exist, or provide your own

```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)

read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```

(FYI: that "\n" you see is a shortcut that tells R to create a new line)

---

## A little extra help

Adding `na = ` to the `read_csv()` command lets you tell R what values should
be interpreted as NAs

```{r}
read_csv("a,b,c\n1,2,.", na = ".")
```

Declaring `quote =` lets you guide R through what it should be reading as
quotation marks (in the case the data doesn't use the typical "")

```{r}
read_delim("x,y\n1,'a,b'", ",", quote = "'")
```

---

## Parsing vectors

The secret sauce behind the `readr` package's ability to read your data is in
its usage of `parse_*()` functions.

```{r}
str(parse_logical(c("TRUE", "FALSE", "NA")))
str(parse_integer(c("1", "2", "3")))
str(parse_date(c("2010-01-01", "1979-10-14")))
parse_integer(c("1", "231", ".", "456"), na = ".")
```

---

## When parsing goes wrong

Parsing failures result in a warning in the console ...

```{r}
x <- parse_integer(c("123", "345", "abc", "123.45"))
```


... and missing data in the output.

```{r}
x
```

Use `problems()` to get a look at all the errors

---

## So. Many. Parsers.

Here's a look at what parsers are available in `readr`

`parse_logical()`

`parse_integer()`

`parse_double()`

`parse_number()`

`parse_character()`

`parse_factor()`

`parse_datetime()`

`parse_date()`

`parse_time()`

Not sure which one you need? Look back at this part of the chapter to decide.

---

## How `readr` works

- It guesses which type of data is in your column based on what it sees in the
first 1000 entries.

- If that first set of entries isn't typical of the column as a whole, it may
guess incorrectly (e.g. If there's nothing but `NA`s in the first 1000 entries,
it'll guess logical)

- Again, you can use `problems()` to get a look at all the errors.

- Best plan of action is to solve issues column by column.

- Once you figure out what type of data the column should be read in as, you
can tell R to read it in that way using the `col_*()` arguments

```{r eval=FALSE, include=TRUE}
challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_double(),
    y = col_logical()
  )
)
```

---

## Other ways to troubleshoot

- Increase the number of observations `readr` uses to guess the column type

```{r eval=FALSE, include=TRUE}
challenge2 <- read_csv(readr_example("challenge.csv"), guess_max = 1001)
```

--

- Nuclear option: Read all the columns in as characters

```{r eval=FALSE, include=TRUE}
challenge2 <- read_csv(readr_example("challenge.csv"), 
  col_types = cols(.default = col_character())
)
```

"This is particularly useful in conjunction with `type_convert()`, which applies the parsing heuristics to the character columns in a data frame."

---

## Writing to a file

Need to get your newly manipulated data out of R? That's where `write_csv()`
and `write_tsv()` come in.

```{r eval=FALSE, include=TRUE}
write_csv(trucks, "trucks.csv")
```

The first argument tells R **which** dataframe we're writing to file. The second
tells it **where** we will write that file.

Planning to use that fresh new csv in Excel? `write_excel_csv()` can make it
play nice with Excel. Also consider checking out the `readxl` package.