
# Examples, Cohort 4 `stringr` package

```{r}
# stringr is a core tidyverse package
library(tidyverse)
```

```{r}
# A simple vector of string values
city <- c("Lincoln", "Houston", "Calgary", "New York City", "Chicago", "Rio de Janeiro")

city
```

# Example 1: Counting length

```{r}
# Notice spaces are counted as characters
str_length(city)
```

# Example 2: Combining strings example

```{r}
homes <- str_c("Lincoln", "Calgary", "Houston")

homes

# Using a seperator
homes <- str_c("Lincoln", "Calgary", "Houston", sep = ", ")

homes
```

# Example 3: Subsetting strings example

```{r}
str_sub(city, 1, 3)

# Start from the end

str_sub(city, -3, -1)
```

# Example 4: Change string case

```{r}
# All upper case
str_to_upper(city)

# All to lower case
str_to_lower(city)
```

# Example 5: Sorting the strings

```{r}
# Sorting strings
str_sort(city)
```

# Example 6: Determine which strings match a pattern

```{r}
# Using str_detect, all cities containing the vowel i
str_detect(city, "[i]")

# Count the number of the vowel i in the city
str_count(city, "[i]")
```

# Example 7: Find positions of matches

```{r}
# Find the length of the string
str_locate(city, "$")

fruit <- c("apple", "banana", "pear", "pineapple")
str_locate(fruit, "e")
```


# Example 8: Extract matches

```{r}
# Extract the first lower-case vowel
str_extract(city, "[aeiou]")

# Extract all the vowels from the string
str_extract_all(city, "[aeiou]")
```


# Example 9: Replace all vowels with a number

```{r}
# The first vowel
str_replace(city, "9", pattern = "[aeiou]")

# All vowels with a 9
str_replace_all(city, "9", pattern = "[aeiou]")
```


# Example 10: Split on match

```{r}
# Split at a space
str_split(city, pattern = "\\s")

# Split the first capitalized letter
str_split(city, pattern = "[A-Z]")

# Split at a vowel
str_split(city, pattern = "[aeiou]")
```


# `stringi` examples

```{r}
library(stringi)

# Reverse the strings
stri_reverse(starwars$name)

# Detect character set and language
stri_enc_detect(starwars$name)

# General stats for a character vector
stri_stats_general(starwars$name)
```


