---
title: "R4DS bookclub"
subtitle: "Chapter 13: Relational Data"
author: "Ruth Alsancak"
date: "05/10/2020"
output:
  xaringan::moon_reader:
    css: [default, rladies, rladies-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

# Relational data

Collectively, multiple tables of data are called **relational data** because it
is the relations, not just the individual datasets, that are important.

A variable in one table and its corresponding variable in another table form a 
**relation**.

<br/>

<img src="https://d33wubrfki0l68.cloudfront.net/245292d1ea724f6c3fd8a92063dcd7b
fb9758d02/5751b/diagrams/relational-nycflights.png" width="60%"/>

---

# Verbs for working with pairs of tables

- Mutating joins

<img src="https://d33wubrfki0l68.cloudfront.net/3abea0b730526c3f053a3838953c35
a0ccbe8980/7f29b/diagrams/join-inner.png" width="50%"/>

--

- Filtering joins

<img src="https://d33wubrfki0l68.cloudfront.net/028065a7f353a932d70d2dfc82bc5c
5966f768ad/85a30/diagrams/join-semi.png" width="50%"/>

--

- Set operations

---

# Working with keys

A **key** is a variable (or set of variables) that uniquely identifies an 
observation:

- A **primary key** uniquely identifies an observation in its own table

- A **foreign key** uniquely identifies an observation in another table

- A **surrogate key** can be added if a table doesn't have a primary key (use 
mutate() and row_number())

--

It's good practice to check your primary keys are unique:

```{r include = FALSE}
library(tidyverse)
library(nycflights13)
```


```{r}
planes %>% 
  count(tailnum) %>% 
  filter(n>1)
```

Duplicate keys in one table may be useful if adding in additional information,
but duplicate keys in both tables usually means there's an error (neither of
the keys uniquely identify an observation).

---

# Defining the key columns to use for the join

Default, **by = NULL**, joins using all variables that appear in both tables 
(natural join):

```{r include = FALSE}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
```

```{r}
flights2 %>% 
  left_join(weather)
```

---

# Defining the key columns to use for the join

**by = "key"** joins using a single variable that has the same name in both 
tables:

```{r}
flights2 %>% 
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")
```

---

# Defining the key columns to use for the join

**by = "x"** (character vector) is like a natural join, but only uses some of
the common variables:

```{r}
flights2 %>% 
  left_join(planes, by = "tailnum")
```

---

# Defining the key columns to use for the join

**by = c("a" = "b")** (named character vector) matches variable a in table x 
to column b in table y:

```{r}
flights2 %>% 
  left_join(airports, c("origin" = "faa"))
```

---

# Mutating joins: inner join

**Inner join**: matches pairs of observations whenever the keys are equal:

<img src="https://d33wubrfki0l68.cloudfront.net/3abea0b730526c3f053a3838953c35a
0ccbe8980/7f29b/diagrams/join-inner.png" width="50%"/>

```{r include = FALSE}
x <- tribble(
  ~key, ~val_x,
    1,  "x1",
    2,  "x2",
    3,  "x3"
)
y <- tribble(
  ~key, ~val_y,
    1,  "y1",
    2,  "y2",
    4,  "y3"
)
```

```{r}
x %>% 
  inner_join(y, by = "key")
```

---

# Mutating joins: outer join

**Outer join**: keeps observations that appear in at least one of 
the tables:

.pull-left[1. A **left join** keeps all observations in x

2. A **right join** keeps all observations in y

3. A **full join** keeps all observations in x and y

```{r include = FALSE}
x <- tribble(
  ~key, ~val_x,
    1,  "x1",
    2,  "x2",
    3,  "x3"
)
y <- tribble(
  ~key, ~val_y,
    1,  "y1",
    2,  "y2",
    4,  "y3"
)
```

```{r}
x %>% 
  full_join(y, by ="key")
```
]

.pull-right[<img src="https://d33wubrfki0l68.cloudfront.net/9c12ca9e12ed26a7
c5d2aa08e36d2ac4fb593f1e/79980/diagrams/join-outer.png" width=85%/>]

---

# Filtering joins: semi-join

A **semi-join** keeps all observations in x that have a match in y:

<img src="https://d33wubrfki0l68.cloudfront.net/028065a7f353a932d70d2dfc82bc5c
5966f768ad/85a30/diagrams/join-semi.png" width="50%"/>

```{r include = FALSE}
top_dest <- flights %>% 
  count(dest, sort = TRUE) %>% 
  head(10)
```

```{r}
flights %>% 
  semi_join(top_dest) %>% 
  head(2)
```

---

# Filtering joins: anti-join

An **anti-join** drops all observations in x that have a match in y:

<img src="https://d33wubrfki0l68.cloudfront.net/f29a85efd53a079cc84c14ba4ba6894
e238c3759/c1408/diagrams/join-anti.png" width="50%"/>

```{r}
flights %>% 
  anti_join(planes, by = "tailnum") %>% 
  count(tailnum, sort = TRUE) %>% 
  head(5)
```

---

# Set operations

**Set operations** expect x and y inputs to have same variables. They work with
complete rows, treat the observations like sets, and compare the values of 
every variable.

--

```{r include = FALSE}
df1 <- tribble(
  ~x, ~y,
  1, 1,
  2, 1
)

df2 <- tribble(
  ~x, ~y,
  1, 1,
  1, 2
)
```

.pull-left[
df1
```{r echo = FALSE}
head(df1) %>% 
  knitr::kable(format = "html", table.attr = "style=width:40%;'")
```
]

.pull-right[
df2
```{r echo = FALSE}
head(df2) %>% 
  knitr::kable(format = "html", table.attr = "style=width:40%;'")
```
]

--

<br/>

**intersect(x, y)** returns only observations in both x and y:

```{r}
intersect(df1, df2)
```

---

# Set operations

**union(x, y)** returns unique observations in x and y:

```{r}
union(df1, df2)
```

<br/>

--

**setdiff(x, y)** returns observations in x, but not in y:

.pull-left[
```{r}
setdiff(df1, df2)
```
]

.pull-right[
```{r}
setdiff(df2, df1)
```
]
