---
title: "R for Data Science"
subtitle: "Chapters 6 & 7: Scripts & EDA"
author: "Pavitra Chakravarty"
institute: "@genomixgmailcom"
date: " `r Sys.Date()` "
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css","custom.css"]
    includes:
      after_body: insert-logo.html
---


```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
library(tidyverse)
library(hexbin)
library(modelr)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```


# Outline

--

* Variation within variables

--

    + variable, value, Observation/Case, Tabular data/Set
    
--

* Covariation between variables

--

* Visualization

---


## VARIATION

--

- Barchart versus Histogram

```{r, echo = FALSE, figures-side_1, fig.show="hold", out.width="50%"}
graphics::par(mfrow=c(1,2))

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)

```

---

## Play around with the bin width

--

```{r, echo = FALSE, figures-side_2, fig.show="hold", out.width="30%"}
graphics::par(mfrow=c(1,3))

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 1.0)

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 1.5)

```

---

## Focus in further with a filter clause

--

```{r, echo = FALSE, figures-side_3}
diamonds %>% 
  filter(carat < 3) %>%
  
ggplot( mapping = aes(x = carat)) +
  geom_histogram(binwidth = 0.01)
```

---

## Are there outliers? Zoom in

--

```{r, echo = FALSE, figures-side_4, fig.show="hold", out.width="50%"}
diamonds %>% 
  filter(carat < 3) %>%
  
ggplot( mapping = aes(x = y)) +
  geom_histogram(binwidth = 0.5)


ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) +
  coord_cartesian(ylim = c(0, 50))
```   


---

## Exercises 

--

### Use ylim to focus further

--

```{r, echo = FALSE, figures-side_5, fig.show="hold", out.width="30%"}
graphics::par(mfrow=c(1,3))

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x), binwidth = 1.0) + 
  coord_cartesian(ylim = c(0, 500))

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = y), binwidth = 1.0) + 
  coord_cartesian(ylim = c(0, 500))

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = z), binwidth = 1.0) + 
  coord_cartesian(ylim = c(0, 500))

```
---

--

### Different bin sizes to see variance better

```{r, echo = FALSE, figures-side_6, fig.show="hold", out.width="33%"}
graphics::par(mfrow=c(1,3))

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(price), binwidth = 1.5) 
  
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(price), binwidth = 5.0) 

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(price), binwidth = 10.0) 
```

---

--

```{r, echo = FALSE, figures-side_7, fig.show="hold", out.width="50%"}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(price), binwidth = 50.0) 

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(price), binwidth = 100.0) 

```

---

--
### Simple count 

```{r,echo=TRUE, echo = FALSE}

diamonds %>% 
  filter(between(carat, 0.99, 3)) %>%
count(carat)

```
---

--
### difference between coord_cartesian and xlim/ylim

```{r, echo = FALSE, figures-side_8, fig.show="hold", out.width="50%"}
graphics::par(mfrow=c(1,3))

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 20) +
  coord_cartesian(xlim = c(0,3000), ylim = c(0,500))

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 20) +
  xlim(c(0,3000)) +
  ylim(c(0,500))
```
---

--

### difference between barplot and histogram

```{r, echo = FALSE, figures-side_17, fig.show="hold", out.width="50%"}
data.frame (a = c(NA, NA, NA, rnorm(1000,0,1))) %>% ggplot() +
  geom_histogram(mapping = aes(x=a), bins= 50)

data.frame (a = c(NA, NA, NA, c('A','A','B','B','B',NA))) %>% ggplot() +
  geom_bar(mapping = aes(x=a))
```
---

--

## COVARIATION

--

### Categorical and Continuous Variable

```{r, echo = FALSE, figures-side_9, fig.show="hold", out.width="50%"}
ggplot(data = diamonds, mapping = aes(x = price)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)

ggplot(data = diamonds, mapping = aes(x = reorder(cut, price, FUN=median), y = price)) +
  geom_boxplot()
```

---

--

```{r, echo = FALSE}

ggplot(data = diamonds, mapping = aes(x = reorder(color, price, FUN=median), y = price)) +
  geom_boxplot()
```

---

--

### 2 categorical variables


```{r, echo = FALSE}
ggplot(data = diamonds, mapping = aes(x = cut, y = color)) +
  geom_count()

 
```


---

--

### 2 continuous variables

```{r, echo = FALSE, figures-side_11, fig.show="hold", out.width="50%"}
diamonds %>% 
  filter(carat < 3) %>%  ggplot() +
  geom_bin2d(mapping = aes(x = carat, y = price))

diamonds %>% 
  filter(carat < 3) %>%  ggplot() +
  geom_hex(mapping = aes(x = carat, y = price))
```

---

--

### 2 continuous variables (one binned)

```{r, echo = FALSE, figures-side_12, fig.show="hold", out.width="50%"}
diamonds %>% 
  filter(carat < 3) %>%  ggplot(mapping = aes(x = carat, y = price)) +
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.1)))

diamonds %>% 
  filter(carat < 3) %>%  ggplot(mapping = aes(x = carat, y = price)) +
  geom_boxplot(mapping = aes(group = cut_number(carat, 20)))
```

---

--

### 2 continuous variables (not binned)

```{r, echo = FALSE, figures-side_13, fig.show="hold", out.width="50%"}
diamonds %>% 
  filter(carat < 3) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = x, y = y)) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))

diamonds %>% 
  filter(carat < 3) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = z, y = y)) +
  coord_cartesian(xlim = c(2.5, 8), ylim = c(4, 11))
```

---

--

### Remove effects of one strongly correlated variable

```{r, echo = FALSE, figures-side_14, fig.show="hold", out.width="50%"}
mod <- lm(log(price) ~ log(carat), data = diamonds)
diamonds2 <- diamonds %>% 
  add_residuals(mod) %>% 
  mutate(resid = exp(resid))
ggplot(data = diamonds2) + 
  geom_point(mapping = aes(x = carat, y = resid))

ggplot(data = diamonds2) + 
  geom_boxplot(mapping = aes(x = cut, y = resid))
```
