---
title: "R for Data Science"
subtitle: "Chapter 4: Workflow basics"
author: "Erick Knackstedt"
institute: "@Eknackstedt"
date: " `r Sys.Date()` "
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css","custom.css"]
    includes:
      after_body: insert-logo.html
---


```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```


# Outline

- Coding basics

--

- What's in a name?

--

- Calling functions

--

- Exercises

---


## Coding basics 

What is a fundamental use for R?

--

- Use R as a calculator

--

What step is a basic building block in programming? 

--

- Create new objects

--

How do you create new objects?

--

```{r eval=FALSE, include=TRUE}
object_name <- value
```

--

> When reading that code say "object name gets value" in your head.

---

## What's in a name?

How do you prefer to write object names? Does it matter?

--

Coding style matters for human reading comprehension. **snake_case** is recommended when an object name has two or more words. 

--

How can you inspect an object's value?

--

By typing its name.

Naming matters for this reason. Keep names descriptive and succinct to ease recall and avoid error. 


---

## Calling functions

How do you call a built in function?

--

```{r eval=FALSE, include=TRUE}
function_name(arg1 = val1, arg2 = val2, ...)
```

--

How does RStudio augment your ability to program?

--

With lots of helpful shortcuts. 

- Start typing and press tab for auto-complete options. 
- ↑/↓ to select an option.
- F1 to access the help page.

--

When you assign a name a value the value is not automatically printed out. Can one call assign and print the value?

--

Yes. Wrap the assignment code in parenthesis. 

```{r}
(y <- seq(1, 10, length.out = 5))
```

---

## Exercises

> 1. Why does this code not work?

```{r eval=FALSE, include=TRUE}
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

--

The object name was misspelled! 

---
### Exercises cont.

> 2. Tweak each of the following R commands so that they run correctly:

--

```{r eval=FALSE, include=TRUE}
ggplot(dota = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

--

```{r eval=FALSE, include=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

--

``` {r eval=FALSE, include=TRUE}
fliter(mpg, cyl = 8)
```

--

```{r eval=FALSE, include=TRUE}
filter(mpg, cyl == 8)
```

--

``` {r eval=FALSE, include=TRUE}
filter(diamond, carat > 3)
```

--

```{r eval=FALSE, include=TRUE}
filter(diamonds, carat > 3)
```

---

### Exercises cont.

> 3. Press Alt + Shift + K. What happenes? How can you get to the same place using the menus?

--

This opens up the keyboard shortcuts help page. You can access this by clicking on Help on the ribbon in the upper left hand corner of RStudio. 

